---
title: "Statistical Inference Part1"
author: "Armin Wallrab"
date: "24. Mai 2015"
output: html_document
---
 
# Statistical Inference 
## Part 1: Simulation

In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. I set lambda = 0.2 for all of the simulations. I'll investigate the distribution of averages of 40 exponentials with a thousand simulations.

```{r}
library(ggplot2)
set.seed(123)
lambda <- 0.2
n <- 40
runs <- 1000
sim <- matrix(rexp(runs*n, rate=lambda), runs)
sim_means <- rowMeans(sim)
```

### 1. Show the sample mean and compare it to the theoretical mean of the distribution

```{r}
# mean of the simulation and the distribution
c(mean(sim_means), 1/lambda)
```

The simulated and theoretical mean values are very close, as expected by the Central Limit Theorem. The distributions are centered around the same values.

### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution

```{r}
# compare the variances
sim_var <- var(sim_means)
the_var <- 1/lambda**2/n
c(sim_var, the_var)
```

The simulated and theoretical variances are also very close.

### 3. Show that the distribution is approximately normal
According to the CLT the distribution of the simulated means should be approximately normal. To illustrate this I'll normalize the vectors and compare it to a standard distribution.
 
```{r}
library(ggplot2)
hist(sim_means, breaks=40, probability=TRUE, main="Distribution of Simulation", xlab="")
lines(density(sim_means), col="green")
abline(v=1/lambda, col="red")
xfit <- seq(min(sim_means), max(sim_means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))
lines(xfit, yfit, pch=22, col="red")
# add a legend
legend('topright', c("Simulated", "Theoretical"), col=c("green", "red"), lty=c(1,1))
```

Also the q-q plot for quantiles shows that the distribution is normal:

```{r}
qqnorm(sim_means, col="green")
qqline(sim_means, col="red")
```

Last but not least, to verify I'll also try to match the confidence intervals which are calculated below:

```{r}
sim_ci <- mean(sim_means) + c(-1,1)*1.96*sd(sim_means)/sqrt(n)
the_ci <- 1/lambda + c(-1,1)*1.96*sqrt(the_var)/sqrt(n);
c(sim_ci, the_ci)
```

This shows that the distribution is normal as expected with the CLT.
